\newcounter{tempcounter}
\setcounter{tempcounter}{\arabic{off}}

\SeqScore s define scoring functions over sequences that can assign a score to each input sequence. Such scores can then be used
to classify sequences by chosing the class represented by that \SeqScore~yielding the maximum score.

Sub-iterfaces of \SeqScore~are \DiffSS, which is tailored to numerical optimization, and \StatMod, which represents statistical models 
defining a proper likelihood. \StatMod, in turn, has two sub-interfaces \TrainSM~and \DiffSM, where the latter joins the properties of \StatMod s and \DiffSS s. All of these will be explained in the sub-sections of this section.

\renewcommand{\codefile}{../../de/jstacs/sequenceScores/SequenceScore.java}
\setcounter{off}{47}

\SeqScore~extends the standard interface \lstinline+Cloneable+ and, hence, implementations must provide a \lstinline+clone()+ method, which returns a deep copy of all fields of an instance:
\code{0}
Since the implementation of the clone method is very score-specific, it must be implemented anew for each implementation of the \SeqScore~interface. Besides that \SeqScore~also extends \Storable, which demands the implementation of the \lstinline+toXML()+ method and a constructor working on a \lstinline+StringBuffer+ containing an XML-representation (see section~\ref{sec:xmlparser}).

\SeqScore~also defines some methods that are basically getters for typical properties of a \SeqScore~implementation. 
These are
\addtocounter{off}{7}
\code{0}
which returns the current \AlphabetContainer~of the model, i.e., the \AlphabetContainer~of \Sequence s this \SeqScore can score,
\addtocounter{off}{7}
\code{0}
which returns a - distinguishing - name of the current \SeqScore~instance, and
\addtocounter{off}{10}
\code{0}
which returns the length of the \SeqScore, i.e. the possible length of input sequences. This length is defined to be zero, if the score can handle \Sequence s of variable length.

The methods
\addtocounter{off}{17}
\code{0}
and
\addtocounter{off}{11}
\code{0}
can be used to return some properties of a \SeqScore~like the number of parameters, the depth of some tree structure, or whatever seems useful. In the latter case, these characteristics are limited to numerical values. If a model is used, e.g., in a cross validation (\KFoldCrossValidation), these numerical properties are averaged over all cross validation iterations and displayed together with the performance measures.

The scoring of input \Sequence is performed by three methods \lstinline+getLogScore+. The name of this method reflects that the returned score should be on a lagorithmic scale. For instance, for a stastical model with a proper likelihood over the sequences this score would correspond to the log-likelihood.
The first method
\addtocounter{off}{10}
\code{0}
returns the score for the given \Sequence. For reasons of efficiency, it is not intended to check if the length or
the \AlphabetContainer~of the \Sequence~match those of the \SeqScore. Hence, no exception is declared for this method. So if
this method is used (instead of the more stringent methods of \StatMod, see \ref{}) this should be checked externally.

The second method is
\addtocounter{off}{13}
\code{0}
which returns the score for the sub-sequence starting at position \lstinline+start+ end extending to the end of the \Sequence or the end of the \SeqScore, whichever comes first.

The third method is
\addtocounter{off}{17}
\code{0}
and computes the score for the sub-sequence between \lstinline+start+ and \lstinline+end+, where the value at position \lstinline+end+ is the last value considered for the computation of the score.

Two additional methods
\addtocounter{off}{25}
\code{0}
and
\addtocounter{off}{26}
\code{0}
allow for the computation of such scores for all \Sequence s in a \DataSet. The first method returns these scores, whereas the second stores the scores to the provided \lstinline+double+ array. This may be reasonable to save memory, for instance if we compute the log-likelihoods of a large number of sequences using different models.

The result of this method (w.r.t. one \Sequence) should be identical
to independent calls of \lstinline+getLogScore+. Hence, these method are pre-implemented exactly in this way in the abstract implementations
of \SeqScore.

Finally, the return value of the method
\addtocounter{off}{9}
\code{0}
indicates if a \SeqScore~has been initialized, i.e., all parameters have been set to some (initial or learned) value and the \SeqScore~is
ready to score \Sequence s.

\subsection{DifferentiableSequenceScores}
\renewcommand{\codefile}{../../de/jstacs/sequenceScores/differentiable/DifferentiableSequenceScore.java}
\setcounter{off}{68}

The interface \DiffSS~extends \SeqScore~ and adds all methods that are necessary for a numerical optimization of the parameters
of a \SeqScore. This especially includes the computation of the gradient w.r.t. the parameters for a given input \Sequence.

Before we can start a numerical optimization, we first need to specify initial parameters. \DiffSS declares two 
methods for that purpos. The first is
\code{1}
which is tailored to initializing the parameters from a given data set. The parameters have the following meaning: \lstinline+index+ is the index of the class this \DiffSS~ represents.
For some parameterizations, we can either optimize (and initialize) only the free parameters of the model, i.e., excluding parameters with values that can be computed from the values of a subset of the remaining parameters, or we can optimize all parameters. For example such parameters could be those of a DNA dice, where the probability of one nucleotide is fixed if we know the probabilities of all other nucleotides.
If \lstinline+freeParams+ is \lstinline+true+, only the free parameters are used.
The array \lstinline+data+ holds the input data, where \lstinline+data[index]+ holds the data for the current class, and \lstinline+weights+ are the weights on each \Sequence~in \lstinline+data+.

The second method for initialization is
\addtocounter{off}{13}
\code{0}
which initializes all parameters randomly, where the exact method for drawing parameter values is implementation-specific.

Like for the method \lstinline+getLogScore+ of \SeqScore, there are three methods for computing the gradient w.r.t. the parameters.
The first is
\addtocounter{off}{21}
\code{0}
where \lstinline+seq+ is the \Sequence~for which the gradient is to be computed, and \lstinline+indices+ and \lstinline+partialDer+ are lists that shall be filled with the indexes and values of the partial derivation, respectively. Assume that we implement a \DiffSS~$f$~with $N$ parameters and $d_i = \frac{\partial \log f}{\partial \lambda_i}$ is the partial derivation of the log-score w.r.t. the $i$-th parameter of $f$. Then we would add the index $i$ to \lstinline+indices+ and (before subsequent additions to \lstinline+indices+ or \lstinline+partialDer+) add $d_i$ to \lstinline+partialDer+. Partial derivations, i.e., components of the gradient, that are zero may be omitted.

The other two methods
\addtocounter{off}{24}
\code{0}
and
\addtocounter{off}{28}
\code{0}
additionally allow for the specification of a start (and end) position within the \Sequence~\lstinline+seq+ in the same manner as specified for \lstinline+getLogScore+ of \SeqScore.

The method
\addtocounter{off}{11}
\code{0}
returns the number of parameters ($N$ in the example above) of parameters of this \DiffSS. It is used, for example, to create temporary arrays for the summation of gradients during numerical optimization.

Since numerical optimization for non-convex problems may get stuck in local optima, we often need to restart numerical optimization multiple times from different initial values of the parameters. As the number of starts that is required to yield the global optimum with high probability is highly score-specific, \DiffSS~defines a method
\addtocounter{off}{8}
\code{0}
that returns the number of starts that is recommended by the developer of this \DiffSS. If multiple \DiffSS~are used in a numerical optimization (for instance for the different classed), the maximum of these values is used in the optimization. If your parameterization is known to be convex, this method should return 1.

The currently set parameters (either by initialization or explicitly, see below) can be obtained by
\addtocounter{off}{16}
\code{0}
and new parameter values may be set by
\addtocounter{off}{12}
\code{0}
where \lstinline+start+ is the index of the value for the first parameter of this \DiffSS~ within \lstinline+params+.

Finally, the method
\addtocounter{off}{12}
\code{0}
returns the class parameter for a given class probability. The default implementation in the abstract class \AbstractDiffSS~returns \lstinline+Math.log(classProb)+. This abstract class also provides standard implementations for many of the other methods.

\subsection{StatisticalModels}
\renewcommand{\codefile}{../../de/jstacs/sequenceScores/statisticalModels/StatisticalModel.java}
\setcounter{off}{122}

\StatMod s extend the specification of \SeqScore s by the computation of proper (normalized) likelihoods for input sequences.
To this end, it defines three methods for computing the log-likelihood of a \Sequence~given a \StatMod~and its current parameter values.

The first method just requires the specification of the \Sequence~object for which the log-likelihood is to be computed:
\code{0}
If the \StatMod~has not been trained prior to calling this method, it is allowed to throw an \lstinline+Exception+. The meaning of this method is slightly different for inhomogeneous, that is position-dependent, and homogeneous models. In case of an inhomogeneous model, for instance a position weight matrix, the specified \Sequence~must be of the same length as the model, i.e. the number of columns in the weight matrix. Otherwise an Exception should be thrown, since users may be tempted to misinterpret the return value as a probability of the complete, possibly longer or shorter, provided sequence.
In case of models that can handle \Sequence s of varying lengths, for instance homogeneous Markov models, this method must return the likelihood of the complete sequence. Since this is not always the desired result, to other methods are specified, which allow for computing the likelihood of sub-sequences. This method should also check if the provided \Sequence~is defined over the same \AlphabetContainer~as the model.

The second method is
\addtocounter{off}{-28}
\code{0}
which computes the log-likelihood of the sub-sequence starting at position \lstinline+startpos+. The resulting value of the log-likelihood must be the same as if the user had called \lstinline+getLogProbFor(sequence.getSubSequence(startpos))+.

The third method reads
\addtocounter{off}{-38}
\code{0}
and computes the likelihood of the sub-sequence starting at position \lstinline+startpos+ up to position \lstinline+endpos+ (inclusive). The resulting value of the likelihood must be the same as if the user had called \lstinline+getProbFor(sequence.getSubSequence(startpos,endpos-startpos+1))+.

If we want to use Bayesian principles for learning the parameters of a model, we need to specify a prior distribution on the parameter values. In some cases, for instance for using MAP estimation in an expectation maximization (EM) algorithm, it is not only necessary to estimate the parameters taking the prior into account, but also to know the value of the prior (or a term proportional to it). For this reason, the \StatMod~interface defines the method
\addtocounter{off}{77}
\code{0}
which returns the logarithm of this prior term.

If the concept of a prior is not applicable for a certain model or other reasons prevent you from implementing this method, \lstinline+getLogPriorTerm()+ should return \lstinline+Double.NEGATIVE_INFINITY+.

\StatMod s can also be used to create new, artificial data according to the model assumptions. For this purpose, the \StatMod~interface specifies a method
\addtocounter{off}{50}
\code{1}
which returns a \DataSet~of \lstinline+numberOfSequences+ \Sequence s drawn from the model using its current parameter values. The second parameter \lstinline+seqLength+ allows for the specification of the lenghts of the drawn \Sequence s. For inhomogeneous model, which inherently define the length of possible sequence, this parameter should be \lstinline+null+ or an array of length 0. For variable-length models, the lengths may either be specified for all drawn sequences by a single \lstinline+int+ value or by an array of length \lstinline+numberOfSequences+ specifying the length of each drawn sequence independently.

The implementation of this method is not always possible. In its default implementation in the abstract implementations of \StatMod, this method just throws an Exception and must be explicitly overridden to be functional.

The last method defined in \StatMod~ is
\addtocounter{off}{11}
\code{0}
which returns the maximum number of preceeding positions that are considered for (conditional) probabilities.
For instance, for a position weight matrix, this method should return 0, whereas for a homogeneous Markov model of order 2, it should return 2.

\subsubsection{TrainableStatisticalModels}

Statistical models that can be learned on a single input data set are represented by the interface \TrainSM~of Jstacs. In most cases, such models are learned by generative learning principles like maximum likelihood (ML) or maximum a-posteriori (MAP). For models that are learned from multiple data sets, commonly by discriminative learning principles, Jstacs provides another interface \DiffSM, which will be presented in the next sub-section.

In the following, we briefly describe all methods that are defined in the \TrainSM~interface. For convenience, an abstract implementation \AbstractTrainSM~of \TrainSM~exists, which provides standard implementations for many of these methods.

\renewcommand{\codefile}{../../de/jstacs/sequenceScores/statisticalModels/trainable/TrainableStatisticalModel.java}
\setcounter{off}{58}

The parameters of statistical models are typically learned from some training data set. For this purpose, \TrainSM~specifies a method \lstinline+train+
\code{0}
that learns the parameters from the training data set \lstinline+data+. By specification, successive calls to \lstinline+train+ must result in a model trained on the data set provided in the last call, as opposed to incremental learning on all data sets.

Besides this simple \lstinline+train+ method, \TrainSM~also declares another method
\addtocounter{off}{27}
\code{0}
that allows for the specification of weights for each input sequence. Since the previous method is a special case of this one where all weights are equal to $1$, only the weighted variant must be implemented, if you decide to extend \AbstractTrainSM. This method should be implemented such that the specification of \lstinline+null+ weights leads to the standard variant with all weights equal to 1. The actual training method may be totally problem and implementation specific. However, in most cases you might want to use one of the generative learning principles ML or MAP.

After a model has been trained it can be used to compute the likelihood of a sequence given the model and its (trained) parameters as defined in \StatMod.

The method
\addtocounter{off}{7}
\code{0}
should return some \lstinline+String+ representation of the current model. Typically, this representation should include the current parameter values in some suitable format.

\renewcommand{\codefile}{../codeExamples/NewCodeExampleTest.java}
\setcounter{off}{\arabic{tempcounter}}

Besides the possibility to implement new statistical models in Jstacs, many of them are already implemented and can readily be used.
As a central facility for creating \TrainSM~instances of many standard models, Jstacs provides a \TrainSMFactory.

Using the \TrainSMFactory, we can create a new position weight matrix (PWM) model by calling
\addtocounter{off}{10}
\code{0}
where we need to specify the (discrete) alphabet, the length of the matrix ($10$), i.e. the number of positions modeled, and an equivalent sample size (ESS) for MAP estimation ($4.0$). For the concept of an equivalent sample size and the BDeu prior used for most models in Jstacs, we refer the reader to (Heckerman). If the ESS is set to $0.0$, the parameters are estimated by the ML instead of the MAP principle.

The factory method for an inhomogeneous Markov model of arbitrary order -- the PWM model is just an inhomogeneous Markov model of order 0 -- is
\addtocounter{off}{1}
\code{0}
where the parameters are in complete analogy to the PWM model, expect the last one specifying the order of the inhomogeneous Markov model, which is 2 in the example.

We can also create permuted Markov models of order $1$ and $2$, where the positions of the sequences may be permuted before building an inhomogeneous Markov model. The permutation is chosen such that the mutual information between adjacent positions is maximized. We create a permuted Markov model of length 7 and order 1 by calling
\addtocounter{off}{1}
\code{0}

Homogeneous Markov models are created by
\addtocounter{off}{1}
\code{0}
where the first parameter again specifies the alphabet, the second parameter is the equivalent sample size, and the third parameter is the order of the Markov model.



\subsubsection{DifferentiableStatisticalModels}

